---
title: "HW7_syQiu"
author: "Shuyi Qiu"
format: html
editor: visual
embed-resources: true
theme: yeti
---

Load packages and dataset

```{r}
#| warning: false
#| error: false
library(gssr)
library(tidyverse)
library(infer)

gss18 <- gss_get_yr(2018) 
```

# 7.1.1 Probabilities

```{r}
d <- gss18 |> 
  select(sex, attend, polviews) |> 
  haven::zap_missing() |> 
  mutate(sex = as_factor(sex)) |> 
  haven::zap_labels() |> 
  drop_na() |> 
  mutate(
    weekly = factor(ifelse(attend >= 7, 1L, 0L), levels = c(0L, 1L), labels = c("LT Weekly", "Weekly")),
    conservative = factor(ifelse(polviews >= 5, 1L, 0L), levels = c(0L, 1L), labels = c("Non-Conservative", "Conservative"))
  )

table(d$weekly, d$conservative)
table(d$weekly)
table(d$conservative)
```

1.  The probability that a person attends religious services weekly.

    $Pr(Weekly)=\frac{497}{2235}=$ `r round(497/2235,4)`

2.  The probability that a person does not attend religious services weekly.

    $Pr(LT\ Weekly)=\frac{1738}{2235}=$ `r round(1738/2235, 4)`

3.  The probability that a person is conservative, given that they attend religious services weekly.

    $Pr(Conservative|Weekly)=\frac{240}{497}=$ `r round(240/497,4)`

4.  The probability that a person attends services weekly, given that they are conservative.

    $Pr(Weekly|Conservative)=\frac{240}{735}=$ `r round(240/735,4)`

5.  The probability that a person is conservative given that they do not got to church weekly

    $Pr(Conservative|LT\ Weekly)=\frac{495}{1738}=$ `r round(495/1738,4)`

6.  The probability that a person is conservative and also attends religious services weekly.

    $Pr(Weekly\&Conservative)=\frac{240}{2235}=$ `r round(240/2235,4)`

7.  The probability that a person is non-conservative but does not attend religious services weekly.

    $Pr(LT\ Weekly\&Non-Conservative)=\frac{1243}{2235}=$ `r round(1243/2235,4)`

8.  The joint probability of W and C, assuming it is independent:

    $Pr(Weekly\&Conservative)=Pr(Weekly)*Pr(Conservative)=\frac{497}{2235}*\frac{735}{2235}=$ `r round((497*735)/2235^2,4)`

# 7.1.2 Table

```{r}
f <- gss18 |> 
  select(happy, marital) |> 
  haven::zap_missing() |> 
  haven::zap_labels() |> 
  drop_na() |> 
  mutate(
    married = factor(ifelse(marital == 1, 1L, 0L), levels = c(0L, 1L), labels = c("Not married", "Married")),
    happyn = factor(ifelse(happy <= 2, 1L, 0L), levels = c(0L, 1L), labels = c("Not happy", "Happy"))
  )

table(f$married, f$happyn)
table(f$married)
table(f$happyn)
```

Redo the exercise in 7.1.1

1.  The probability that a person is happy.

    $Pr(happy)=\frac{2006}{2342}=$ `r round(2006/2342,4)`

2.  The probability that a person is not happy.

    $Pr(not\ happy)=\frac{336}{2342}=$ `r round(336/2342,4)`

3.  The probability that a person is married, given that they are happy.

    $Pr(married|happy)=\frac{936}{2006}=$ `r round(936/2206,4)`

4.  The probability that a person is happy, given that they are married.

    $Pr(happy|married)=\frac{936}{997}=$ `r round(936/997,4)`

5.  The probability that a person is married, given that they are not happy.

    $Pr(married|not\ happy)=\frac{61}{336}=$ `r round(61/336,4)`

6.  The probability that a person is both happy and married.

    $Pr(happy\&married)=\frac{936}{2342}=$ `r round(936/2342,4)`

7.  The probability that a person is not happy and not married.

    $Pr(not\ happy\&not\ married)=\frac{275}{2342}=$ `r round(275/2342,4)`

8.  The joint probability that a person is happy and married, given that they are independent.

    $Pr(happy\&married)=Pr(happy)*Pr(married)=\frac{2006}{2342}*\frac{997}{2342}=$ `r round((2006*997)/2342^2,4)`

# 7.2.1 Four Dependence

## 1. Difference in Probabilities

$Pr(Weekly|Conservative)-Pr(Weekly|Non-conservative)=\frac{240}{735}-\frac{257}{1500}=$ `r round(240/735-257/1500,4)`

## 2. Relative Risk Ratio

$\frac{Pr(Weekly|Conservative)}{Pr(Weekly|Non-conservative)}=\frac{\frac{240}{735}}{\frac{257}{1500}}=$ `r round((240/735)/(257/1500),4)`

## 3. Odds Ratio

Odds of that a conservative person goes to religion services weekly = $\frac{240}{495}=$ `r round(240/495,4)`

Odds of that a non-conservative person goes to religion services weekly = $\frac{257}{1243}=$ `r round(257/1243,4)`

Odds ratio = $\frac{\frac{240}{495}}{\frac{257}{1243}}=$ `r round((240/495)/(257/1243),4)`

## 4. Log Odds Ratio

$log(Odds\ Ratio)=log(\frac{\frac{240}{495}}{\frac{257}{1243}})=$ `r round(log((240/495)/(257/1243)),4)`

# 7.2.2. Inference

## 1. Relative Risk Ratio

```{r}
boot_dist_d1 <- d |> 
  specify(weekly ~ conservative, success = "Weekly") |> 
  generate(
    reps = 1000, # Do 1000 times
    type = "bootstrap" # Draw sample where the sample size equals tot the input sample size with replacement
  ) |> 
  calculate(stat = "ratio of props",
            order = c("Conservative","Non-Conservative"))

ci <- boot_dist_d1 |> 
  get_confidence_interval(level = 0.95)

boot_dist_d1 |> 
  visualise() +
  shade_ci(ci)
```

## 2. Odds Ratio

```{r}
boot_dist_d2 <- d |> 
  specify(weekly ~ conservative, success = "Weekly") |> 
  generate(
    reps = 1000, # Do 1000 times
    type = "bootstrap" # Draw sample where the sample size equals tot the input sample size with replacement
  ) |> 
  calculate(stat = "odds ratio",
            order = c("Conservative","Non-Conservative"))

ci <- boot_dist_d2 |> 
  get_confidence_interval(level = 0.95)

boot_dist_d2 |> 
  visualise() +
  shade_ci(ci)
```

# 7.2.3 Four Dependence - My own example

## 1. Difference in Probabilities

$Pr(Happy|Married)-Pr(Happy|Not married)=\frac{936}{997}-\frac{1070}{1345}=$ `r round(936/997-1070/1345,4)`

## 2. Relative Risk Ratio

$Pr(Happy|Married)-Pr(Happy|Not married)=\frac{\frac{936}{997}}{\frac{1070}{1345}}=$ `r round((936/997)/(1070/1345),4)`

## 3. Odds Ratio

The odds that a married person is happy = $\frac{936}{61}=$ `r round(936/61,4)`

The odds that a person who's not married is happy = $\frac{1070}{275}=$ `r round(1070/275,4)`

The odds ratio = $\frac{\frac{936}{61}}{\frac{1070}{275}}=$ `r round((936/61)/(1070/275),4)`

## 4. Log Odds Ratio

Log odds ratio = $log(\frac{\frac{936}{61}}{\frac{1070}{275}})=$ `r round(log((936/61)/(1070/275)),4)`

# 7.2.4 Sample Distribution - My own example

## 1. Relative Risk Ratio

```{r}
boot_dist_f1 <- f |> 
  specify(happyn ~ married, success = "Happy") |> 
  generate(
    reps = 1000, # Do 1000 times
    type = "bootstrap" # Draw sample where the sample size equals tot the input sample size with replacement
  ) |> 
  calculate(stat = "ratio of props",
            order = c("Married","Not married"))

ci <- boot_dist_f1 |> 
  get_confidence_interval(level = 0.95)

boot_dist_f1 |> 
  visualise() +
  shade_ci(ci)
```

## 2. Odds Ratio

```{r}
boot_dist_f2 <- f |> 
  specify(happyn ~ married, success = "Happy") |> 
  generate(
    reps = 1000, # Do 1000 times
    type = "bootstrap" # Draw sample where the sample size equals tot the input sample size with replacement
  ) |> 
  calculate(stat = "odds ratio",
            order = c("Married","Not married"))

ci <- boot_dist_f2 |> 
  get_confidence_interval(level = 0.95)

boot_dist_f2 |> 
  visualise() +
  shade_ci(ci)
```
