---
title: "HW5_syQiu"
author: "Shuyi Qiu"
format: html
editor: visual
---

```{r}
#| warning: false
#| error: false
library(tidyverse)
```

Some basic functions to work with binomial distribution:

-   `dbinom(x, size, prob)`: return the probability of having `x` heads if toss a coin for `size` times with a probability of `prob` to have heads. The result is equal to `choose(size, x) * prob^x * (1-prob)^(size - x)`

-   `pbinom(q, size, prob)`: return the area (CDF). `pbinom(size, size, prob) = 1`, not so commonly used...cannot think of a context to use it.

-   `rbinom(n, size, prob)`: randomly generate a vector of `n` random variables, each of which equal to the number of heads within `size` trials of coin toss.

-   `qbinom(p, size, prob)`: inverse of `pbinom()`, return the number at the point when area under the CDF equal to `p`

# 5.2.1 Bernoulli Distribution

```{r}
# Verification of equation 5.1
dbinom(1, 1, 2/3)
dbinom(0, 1, 2/3)
dbinom(2, 1, 2/3)

# Verification of equation 5.2
pbinom(-1, 1, 2/3)
pbinom(0, 1, 2/3)
pbinom(1, 1, 2/3)
pbinom(2, 1, 2/3)

# rbinom() and plot
bernoulli_sample <- data.frame(
  sid = c(1:100),
  results = rbinom(100, 1, 2/3)
)
ggplot(bernoulli_sample, aes(results)) +
  geom_bar()
```

# 5.3.1 Sample space

Possible scenarios by the number of heads:

1 possible way to get 0 $H$: $(T,T,T,T)$

4 possible ways to get 1 $H$: $(H, T, T, T), (T, H, T, T), (T, T, H, T), (T, T, T, H)$

6 possible ways to get 2 $H$s: $(T, T, H, H), (T, H, H, T), (H, H, T, T), (T, H, T, H), (H, T, H, T), (H, T, T, H)$

4 possible ways to get 3 $H$s: $(T, H, H, H), (H, T, H, H), (H, H, T, H), (H, H, H, T)$

1 possible way to get 4 $H$s: $(H, H, H,H)$

Therefore, the sample space of a total of 16 possible ways is

$$
\Omega=\{(T, T, T, T), (H, T, T, T), (T, H, T, T), (T, T, H, T), \\(T, T, T, H), (T, T, H, H), (T, H, H, T), (H, H, T, T), \\(T, H, T, H), (H, T, H, T), (H, T, T, H), (T, H, H, H), \\(H, T, H, H), (H, H, T, H), (H, H, H, T), (H, H, H, H)\}
$$

# 5.3.2 Probabilities

Denote $Y()$ as the number of events

$Y(X=0)=1$

$Y(X=1)=4$

$Y(X=2)=6$

$Y(X=3)=4$

$Y(X=4)=1$

$Pr(X=2)=\frac{Y(X=1)}{\sum_{i=0}^{4}Y(X=i)}=\frac{6}{1+4+6+4+1}=0.375$

$Pr(X<=1)=\frac{Y(X=0)+Y(X=1)}{\sum_{i=0}^{4}Y(X=i)}=\frac{1+4}{1+4+6+4+1}=0.3125$

```{r}
dbinom(2, 4, 0.5) # Pr(x = 2)
pbinom(1, 4, 0.5) # Pr(X <= 1)
```

# 5.3.3 Size = 5

```{r}
binom_sample <- data.frame(
  results = rbinom(n = 1000, size = 5, prob = 2/3)
)
ggplot(binom_sample, aes(x = results)) +
  geom_bar() +
  scale_x_continuous(breaks = c(0:5))

mean(binom_sample$results %% 2 == 0) # Prob(X is an even number)
dbinom(x = 2, size = 5, prob = 2/3) + dbinom(x = 4, size = 5, prob = 2/3) # Theoretically
```

# 5.3.5 Revisit

Calculate manually:

There are 6 possible scenarios to get two heads, therefore, the probability of having two heads in four coin tosses when the probability of head is 1/3 is $Pr(X=2)=6*(\frac{1}{3})^2*(\frac{2}{3})^2=$ `r choose(4,2)*(1/3)^2*(2/3)^2`

```{r}
dbinom(2, 4, 1/3) # Pr(x = 2)
```

# 5.3.7 Verification

We can verify this by creating a data frame listing all the probabilities and cumulative probabilities, as well as the computed results by equation manually.

```{r}
binom_verif <- data.frame(x = c(1:15)) |> 
  mutate(
    func_dbinom = dbinom(x, 15, 2/3),
    calc_dbinom = choose(15, x) * (2/3)^x * (1-2/3)^(15-x),
    func_pbinom = pbinom(x, 15, 2/3),
    calc_pbimom = cumsum(calc_dbinom)
  )

binom_verif
```

We can see the results computed by function and by hand are almost the same.

There are very subtle differences between the cumulative results computed by function and by hand, possibly caused by digits issue.

# 5.4.1 Likelihoods

```{r}
grid <- tibble(prob = seq(0, 1, by = 0.001))
grid$like <- dbinom(21, 47, grid$prob) 
grid <- grid |> 
  mutate(
    clike_raw = cumsum(like),
    clike_normalized = clike_raw/sum(like)
  )

ggplot(grid, aes(x = prob, y = clike_normalized)) +
  geom_line()

1-grid$clike_normalized[grid$prob == 0.588]
```

The chances that `prob` is equal to or greater than 0.588 is 0.025
