---
title: "HW8_syQiu"
format: html
editor: visual
embed-resources: true
---

Set up packages and import data

```{r}
#| warning: false
#| error: false
library(gssr)
library(tidyverse)
library(infer)
library(janitor)

gss18 <- gss_get_yr(2018) 
```

For easy look-up, I listed the definition for each variable here before the exploratory data analyses.

-   `wrkslf`: Whether self-employed or not

-   `attend`: The frequency of attending religious services

-   `polviews` : Political view

-   `cappun`: Favor death penalty or not

-   `suicide3`: Do you agree that a person has the right to suicide if this person has dishonored his/her their?

-   `xmarsex`: Attitude towards extra-marital sexual relationship

-   `degree`: Highest degree achieved

-   `padeg`: R's father's highest degree

-   `madeg`: R's mother's highest degree

-   `sex`: R's sex

-   `race`: R's race

-   `born`: Were you born in US?

-   `mcsds6`: Like to gossip at times

-   `condom`: Used condom or not the last time when the R had sex

-   `marital`: Marital status

-   `god`: Belief in god

# 1. Parents' Education v.s. R's Education

I am interested in the association between parents' education outcome and the child's education outcome. Here I specifically looked into the outcome of college completion.

```{r}
tempdf <- gss18 |> 
  select(degree, padeg, madeg) |> 
  drop_na() |> 
  mutate(
    r_gtba = factor(ifelse(degree >= 3, TRUE, FALSE), levels = c(FALSE, TRUE), labels = c("No BA", "BA or more")),
    f_gtba = factor(ifelse(padeg >= 3, TRUE, FALSE), levels = c(FALSE, TRUE), labels = c("No BA", "BA or more")),
    m_gtba = factor(ifelse(madeg >= 3, TRUE, FALSE), levels = c(FALSE, TRUE), labels = c("No BA", "BA or more"))
  )
```

The three education outcome variables used are defined as `0`: No BA degree, `1`: BA degree or more.

-   `r_gtba`: Whether respondent have a degree of BA or more

-   `f_gtba`: Whether respondent's father's have a degree of BA or more

-   `m_gtba`: Whether respondent's mother's have a degree of BA or more

## 1.1 Father's Education

I looked into the association between father's having a degree of BA or more and the respondent's having a degree of BA or more in this section.

$Y$: Respondent's having BA degree or more

$X$: Respondent's father having BA degree or more

```{r}
tempdf |> 
  tabyl(f_gtba, r_gtba) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(f_gtba, r_gtba) |> 
  adorn_percentages(denominator = "row")
```

Linear probability model: $p(RespBA = 1) = \alpha + \beta FatherBA$

$$
\alpha=Pr(RespBA =1|FatherBA=0)=0.2818
$$

$$
\begin{aligned}
\beta&=Pr(RespBA =1|FatherBA=1)-Pr(RespBA =1|FatherBA=0)\\
&=0.6455-0.2818 \\
&=0.3638
\end{aligned}
$$

Interpretation of $\beta$ : The respondents whose father has BA degree or more have a higher probability of having BA degree or more than those whose father does not have BA degree by 36.38 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta FatherBA$

Where $p$ is $Pr(Resp=1)$

$$
\alpha = log(\frac{366}{933})=-0.9358
$$

$$
\beta=log(\frac{224}{123})-log(\frac{366}{933})=1.5352
$$

Interpretation of $\beta$: The respondents whose father has BA degree or more have a higher log odds of having BA degree or more than those whose father does not have BA degree by 1.54.

## 1.2 Mother's Education

Similarly, I looked into the association between mother's having a degree of BA or more and the respondent's having a degree of BA or more in this section.

$Y$: Respondent's having BA degree or more

$X$: Respondent's mother having BA degree or more

```{r}
tempdf |> 
  tabyl(m_gtba, r_gtba) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(m_gtba, r_gtba) |> 
  adorn_percentages(denominator = "row")
```

Linear probability model: $p(RespBA = 1) = \alpha + \beta MotherBA$

$$
\alpha=Pr(RespBA =1|MotherBA=0)=0.2891
$$

$$
\begin{aligned}
\beta&=Pr(RespBA =1|MotherBA=1)-Pr(RespBA =1|MotherBA=0)\\
&=0.6559-0.2891 \\
&=0.3668
\end{aligned}
$$

Interpretation of $\beta$ : The respondents whose mother has BA degree or more have a higher probability of having BA degree or more than those whose mother does not have BA degree by 36.68 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta MotherBA$

Where $p$ is $Pr(Resp=1)$

$$
\alpha = log(\frac{386}{949})=-0.8996
$$

$$
\beta=log(\frac{204}{107})-log(\frac{386}{949})=1.5449
$$

Interpretation of $\beta$: The respondents whose mother has BA degree or more have a higher log odds of having BA degree or more than those whose mother does not have BA degree by 1.54.

(Without any further adjustment, the crude logistic regression model shows that the effect size of father's having a BA degree or more is almost the same as the effect size of mother's having a BA degree or more on child's education outcome)

# 2. Gender Difference in Attitudes on Death Penalty

```{r}
tempdf <- gss18 |> 
  select(sex, cappun) |> 
  drop_na() |> 
  mutate(
    r_male = factor(ifelse(sex == 1, T, F), levels = c(F, T), labels = c("Female","Male")),
    r_favor = factor(ifelse(cappun == 1, T, F), levels = c(F, T), labels = c("Oppose","Favor"))
  )
```

In this section, I am interested in the association between sex and people's attitude towards the death penalty for persons convicted of murder. The variables are defined as below:

$Y$: Respondent's attitude towards death penalty for persons convicted of murder. 1 if favor, 0 if oppose.

$X$: Respondent's sex. 1 if male, 0 if female.

```{r}
tempdf |> 
  tabyl(r_male, r_favor) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(r_male, r_favor) |> 
  adorn_percentages(denominator = "row")
```

Linear probability model: $p(Favor = 1) = \alpha + \beta Male$

$$
\alpha=Pr(Favor =1|Male=0)=0.5963
$$

$$
\begin{aligned}
\beta&=Pr(Favor =1|Male=1)-Pr(Favor =1|Male=0)\\
&=0.6729-0.5963 \\
&=0.0767
\end{aligned}
$$

Interpretation of $\beta$ : Male respondents have a higher probability of favoring death penalty for persons convicted of murder than female respondents by 7.67 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta Male$

Where $p$ is $Pr(Favor=1)$

$$
\alpha = log(\frac{706}{478})=0.3900
$$

$$
\beta=log(\frac{679}{330})-log(\frac{706}{478})=0.33
$$

Interpretation of $\beta$: Male respondents have a higher log odds of favoring death penalty for persons convicted of murder than female respondents by 0.33.

# 3. Gender Difference in Attitudes Towards Suicide

```{r}
tempdf <- gss18 |> 
  select(sex, suicide3) |> 
  drop_na() |> 
  mutate(
    r_male = factor(ifelse(sex == 1, T, F), levels = c(F, T), labels = c("Female","Male")),
    r_scd = factor(ifelse(suicide3 == 1, T, F), levels = c(F, T), labels = c("No","Yes"))
  )
```

In this section, I am interested in the association between sex and people' attitudes toward people's right to suicide if this person dishonored their families. The variables are defined as below:

$Y$: Whether the respondent think a person has the right to end his or her own life if this person dishonored his or her family. 1 if the respondent agree, 0 if not.

$X$: Respondent's sex. 1 if male, 0 if female.

```{r}
tempdf |> 
  tabyl(r_male, r_scd) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(r_male, r_scd) |> 
  adorn_percentages(denominator = "row")
```

Linear probability model: $p(Suicide = 1) = \alpha + \beta Male$

$$
\alpha=Pr(Suicide =1|Male=0)=0.1144
$$

$$
\begin{aligned}
\beta&=Pr(Suicide =1|Male=1)-Pr(Suicide =1|Male=0)\\
&=0.1408-0.1144 \\
&=0.0264
\end{aligned}
$$

Interpretation of $\beta$ : Male respondents have a higher probability of agreeing that the person who dishonored their families has the right to suicide than female respondents by 2.64 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta Male$

Where $p$ is $Pr(Suicide=1)$

$$
\alpha = log(\frac{98}{759})=-2.0470
$$

$$
\beta=log(\frac{96}{586})-log(\frac{98}{759})=0.2381
$$

Interpretation of $\beta$: Male respondents have a higher log odds of agreeing that the person who dishonored their families has the right to suicide than female respondents by 0.24.

# 4. Girls like gossips?

```{r}
tempdf <- gss18 |> 
  select(sex, mcsds6) |> 
  drop_na() |> 
  mutate(
    r_male = factor(ifelse(sex == 1, T, F), levels = c(F, T), labels = c("Female","Male")),
    r_gsp = factor(ifelse(mcsds6 == 1, T, F), levels = c(F, T), labels = c("No","Yes"))
  )
```

In this section, I am interested in the association between sex and people' self-recognition as a person like gossips at times. The variables are defined as below:

$Y$: Whether the respondent think themselves like to gossip at times. 1 if the respondent think so, 0 if not.

$X$: Respondent's sex. 1 if male, 0 if female.

```{r}
tempdf |> 
  tabyl(r_male, r_gsp) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(r_male, r_gsp) |> 
  adorn_percentages(denominator = "row")
```

Linear probability model: $p(Gossip = 1) = \alpha + \beta Male$

$$
\alpha=Pr(Gossip =1|Male=0)=0.5806
$$

$$
\begin{aligned}
\beta&=Pr(Gossip =1|Male=1)-Pr(Gossip =1|Male=0)\\
&=0.4177-0.5806 \\
&=-0.1629
\end{aligned}
$$

Interpretation of $\beta$ : Male respondents have a lower probability of considering themselves like to gossip at times than female respondents by 16.29 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta Male$

Where $p$ is $Pr(Gossip=1)$

$$
\alpha = log(\frac{342}{247})=0.3254
$$

$$
\beta=log(\frac{231}{322})-log(\frac{342}{247})=-0.6576
$$

Interpretation of $\beta$: Male respondents have a lower log odds of considering themselves like to gossip at times than female respondents by 0.66.

# 5. Religion and Suicide

```{r}
tempdf <- gss18 |> 
  select(attend, suicide3) |> 
  drop_na() |> 
  mutate(
    r_weekly = factor(ifelse(attend >= 6, T, F), levels = c(F, T), labels = c("LT Weekly","Weekly")),
    r_scd = factor(ifelse(suicide3 == 1, T, F), levels = c(F, T), labels = c("No","Yes"))
  )
```

In this section, I am interested in the association between people's frequency of attending religion services and their attitudes towards suicide if the people dishonored their families. The variables are defined as below:

$Y$: Whether the respondent agree that a person has the right to end his or her own life if this person dishonored his or her family. 1 if agree, 0 if not.

$X$: Respondent's frequency of attending religious services. 1 if weekly or more, 0 if less than weekly.

```{r}
tempdf |> 
  tabyl(r_weekly, r_scd) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(r_weekly, r_scd) |> 
  adorn_percentages(denominator = "row")
```

Linear probability model: $p(Suicide = 1) = \alpha + \beta Weekly$

$$
\alpha=Pr(Suicide =1|Weekly=0)=0.1500
$$

$$
\begin{aligned}
\beta&=Pr(Suicide =1|Weekly=1)-Pr(Suicide =1|Weekly=0)\\
&=0.0620-0.1500 \\
&=-0.0879
\end{aligned}
$$

Interpretation of $\beta$ : People who attend the religious services more than weekly have a lower probability of agreeing on suicides because of family dishonor than people who attend religious services less than weekly by 8.79 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta Weekly$

Where $p$ is $Pr(Suicide=1)$

$$
\alpha = log(\frac{169}{958})=-1.7349
$$

$$
\beta=log(\frac{25}{378})-log(\frac{169}{958})=-0.9811
$$

Interpretation of $\beta$: People who attend the religious services more than weekly have a lower log odds of agreeing on suicides because of family dishonor than people who attend religious services less than weekly by 0.98.

# 6. Education and Belief in God

```{r}
tempdf <- gss18 |> 
  select(degree, god) |> 
  drop_na() |> 
  mutate(
    r_gtba = factor(ifelse(degree >= 3, T, F), levels = c(F, T), labels = c("No BA","BA or more")),
    r_god = factor(ifelse(god == 6, T, F), levels = c(F, T), labels = c("Not so believe","Believe without doubts"))
  )
```

In this section, I am interested in the association between people's belief in god and their education level. The variables are defined as below:

$Y$: Whether the respondent believe in god without doubts. 1 if so, 0 if not.

$X$: Whether the respondent have a bachelor's degree or more. 1 if so, 0 if not.

```{r}
tempdf |> 
  tabyl(r_gtba, r_god) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(r_gtba, r_god) |> 
  adorn_percentages(denominator = "row")
```

Linear probability model: $p(God = 1) = \alpha + \beta BA$

$$
\alpha=Pr(God =1|BA=0)=0.5716
$$

$$
\begin{aligned}
\beta&=Pr(God =1|BA=1)-Pr(God =1|BA=0)\\
&=0.4830-0.5716 \\
&=-0.0887
\end{aligned}
$$

Interpretation of $\beta$ : People who have a bechelor's degree or more have a lower probability of believing in god without any doubts than those who do not have a BA degree by 8.87 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta BA$

Where $p$ is $Pr(God=1)$

$$
\alpha = log(\frac{922}{691})=0.2884
$$

$$
\beta=log(\frac{340}{364})-log(\frac{922}{691})=-0.3566
$$

Interpretation of $\beta$: People who have a bachelor's degree or more have a lower log odds of believing in god without any doubts than those who do not have a BA degree by 0.36.

# 7. Birthplace and Education

This section will look into the association between US-born and education level. I will cover two education outcomes: whether completed high school or not, and whether having bachelor's degree or not.

```{r}
tempdf <- gss18 |> 
  select(degree, born) |> 
  drop_na() |> 
  mutate(
    r_gths = factor(ifelse(degree >= 1, T, F), levels = c(F, T), labels = c("No HS","HS or more")),
    r_gtba = factor(ifelse(degree >= 3, T, F), levels = c(F, T), labels = c("No BA","BA or more")),
    r_us = factor(ifelse(born == 1, T, F), levels = c(F, T), labels = c("Not US-born","US-born"))
  )
```

## 7.1 High School Diploma

```{r}
tempdf |> 
  tabyl(r_us, r_gths) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(r_us, r_gths) |> 
  adorn_percentages(denominator = "row")
```

$Y$: Whether the respondent have high school diploma or not. 1 if so, 0 if not.

$X$: Whether the respondent was born in US. 1 if so, 0 if not.

Linear probability model: $p(HS = 1) = \alpha + \beta USborn$

$$
\alpha=Pr(HS =1|USborn=0)=0.7450
$$

$$
\begin{aligned}
\beta&=Pr(HS =1|USBorn=1)-Pr(HS =1|USBorn=0)\\
&=0.9095-0.7450 \\
&=0.1645
\end{aligned}
$$

Interpretation of $\beta$ : People who was born in US have a higher probability to have high school diploma than those who was not born in US by 16.45 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta USBorn$

Where $p$ is $Pr(HS=1)$

$$
\alpha = log(\frac{225}{77})=1.0723
$$

$$
\beta=log(\frac{1860}{185})-log(\frac{225}{77})=1.2357
$$

Interpretation of $\beta$: People who was born in US have a higher log odds to have high school diploma than those who was not born in US by 1.24.

## 7.2 Bachelor's Degree

```{r}
tempdf |> 
  tabyl(r_us, r_gtba) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(r_us, r_gtba) |> 
  adorn_percentages(denominator = "row")
```

$Y$: Whether the respondent have bachelor's degree or not. 1 if so, 0 if not.

$X$: Whether the respondent was born in US. 1 if so, 0 if not.

Linear probability model: $p(BA = 1) = \alpha + \beta USborn$

$$
\alpha=Pr(BA =1|USborn=0)=0.3079
$$

$$
\begin{aligned}
\beta&=Pr(BA =1|USBorn=1)-Pr(BA =1|USBorn=0)\\
&=0.3027-0.3079 \\
&=-0.0053
\end{aligned}
$$

Interpretation of $\beta$ : People who was born in US have a lower probability to have BA degree or more than those who was not born in US by 0.53 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta USBorn$

Where $p$ is $Pr(HS=1)$

$$
\alpha = log(\frac{93}{209})=-0.8097
$$

$$
\beta=log(\frac{619}{1426})-log(\frac{93}{209})=-0.0248
$$

Interpretation of $\beta$: People who was born in US have a lower log odds to have BA degree or more than those who was not born in US by 0.02.

Interesting finding from these two studies in section 7: Whether US-born or not have a strong association with high school diploma, but not a strong one with BA degree. Further information is needed (when the respondent moved to US) to explore the association between education level and birthplace.

# 8. Belief in God and Extra-marital Sex

```{r}
tempdf <- gss18 |> 
  select(god, xmarsex) |> 
  drop_na() |> 
  mutate(
    r_exmasex = factor(ifelse(xmarsex == 1, T, F), levels = c(F, T), labels = c("Not always","Always wrong")),
    r_god = factor(ifelse(god == 6, T, F), levels = c(F, T), labels = c("Not so believe","Believe without doubts"))
  )
```

In this section, I am interested in the association between people's belief in god and their attitudes towards extra-marital sex. The variables are defined as below:

$Y$: Whether the respondent believe in god without doubts. 1 if so, 0 if not.

$X$: Whether the respondent always think extra-marital sex wrong. 1 if so, 0 if not.

```{r}
tempdf |> 
  tabyl(r_god, r_exmasex) |> 
  adorn_totals(c("row","col"))

tempdf |> 
  tabyl(r_god, r_exmasex) |> 
  adorn_percentages(denominator = "row")
```

Linear probability model: $p(ExtraMarSex = 1) = \alpha + \beta God$

$$
\alpha=Pr(ExtraMarSex =1|God=0)=0.5930
$$

$$
\begin{aligned}
\beta&=Pr(ExtraMarSex =1|God=1)-Pr(ExtraMarSex =1|God=0)\\
&=0.8852-0.5930 \\
&=0.2922
\end{aligned}
$$

Interpretation of $\beta$ : People who was believe in god without any doubts have a higher probability of always considering extra-marital sex as wrong than those who do not believe in god so completely by 29.22 percentage points.

Logistic regression model: $logit(\frac{p}{1-p})=\alpha + \beta God$

Where $p$ is $Pr(ExtraMarSex=1)$

$$
\alpha = log(\frac{408}{280})=0.3765
$$

$$
\beta=log(\frac{748}{97})-log(\frac{408}{280})=1.6662
$$

Interpretation of $\beta$: People who was believe in god without any doubts have a higher log odds of always considering extra-marital sex as wrong than that of those who do not believe in god so completely by 1.67.
