---
title: "HW11_syQiu"
author: "Shuyi Qiu"
format: html
editor: visual
embed-resources: true
---

Set-up

```{r}
#| warning: FALSE
#| error: FALSE
library(tidyverse)
theme_set(theme_light(base_family = "Optima"))

library(modelsummary)
library(broom)
library(gt)
library(performance)

library(gssr)
gss18 <- gss_get_yr(2018) 

vars <- c(
  "hompop", "sibs", "numwomen", "nummen", "age", "sex", "race", "attend", "polviews", "degree", 
  "happy", "sexornt", "premarsx", "condom", "wrkslf", "fefam", "cappun", "padeg", "madeg"
)

d <- gss18 |> 
  select(all_of(vars)) |> 
  haven::zap_missing() |> 
  ## continuous vars 
  mutate(across(all_of(vars[1:5]), haven::zap_labels)) |> 
  ## categorical vars
  #mutate(across(!vars[1:5], haven::as_factor)) |>
  mutate(numtot = numwomen + nummen)
```

# 11.1 Goodness of Fit

## 11.1.1 Model Fit

```{r}
df <- d |> 
  mutate(
    rba = ifelse(degree >= 3, T, F),
    sibs = as.numeric(sibs)
    ) |> 
  select(rba, padeg, madeg, sibs) |> 
  drop_na()

logit_mod1 <- glm(formula = rba ~ padeg + madeg + sibs, data = df, family = binomial(link = "logit"))

summary(logit_mod1)

performance_hosmer(logit_mod1)
```

The Hosmer-Lemeshow test compares the observed and predicted proportions within each group. Therefore, a large value of Chi-squared with small p-value shows that there is a significant gap between the observed value and predicted value, indicating a poor fit of the model. A low value of Chi-squared with a large p-value shows that the predicted value and observed value is very close, therefore the model is a good fit.

## 11.1.2 Link Test

```{r}
df$pred_rba = predict(logit_mod1)
logit_mod1_link = glm(formula = rba ~ pred_rba, data = df, family = binomial(link = "logit"))

summary(logit_mod1_link)
```

The $\beta$ coefficient is 1. It means that the model is a good fit.

# 11.2 Poisson Regression

## 11.2.1 Regression & Plot

```{r}
lambda_value <- c(1, 4, 10, 20, 50)
draw_num <- 1000
pois_mat <- matrix(NA, ncol = 2)
for (i in c(1:length(lambda_value))){
  temp <- matrix(NA, nrow = draw_num, ncol = 2)
  temp[,1] <- rep(lambda_value[i], draw_num)
  temp[,2] <- rpois(n = draw_num, lambda = lambda_value[i])
  pois_mat <- rbind(pois_mat, temp)
}

# remove the start value
pois_mat <- pois_mat[c(2:nrow(pois_mat)),]

pois_df <- as.data.frame(pois_mat)
colnames(pois_df) <- c("lambda", "value")

# Plot
ggplot(pois_df |> mutate(lambda = as.factor(lambda)), aes(x = value, group = lambda, fill = lambda)) +
  geom_bar(alpha = 0.8)
```

It seems that when $\lambda$ value is larger and larger, the shape of a poisson distribution is more and more close to a normal distribution with a mean equal to $\lambda$

## 11.2.2 `numtot`

```{r}
df <- d |> 
  mutate(
    sex_new = as.factor(ifelse(sex == 1, "male","female")),
    sex_new = relevel(sex_new, ref = "female"),
    sexornt_new = as.factor(case_when(
      sexornt == 1 ~ "homo",
      sexornt == 2 ~ "bi",
      sexornt == 3 ~ "str"
    )),
    sexornt_new = relevel(sexornt_new, ref = "str"),
    polviews_new = as.factor(case_when(
      polviews <= 3 ~ "liberal",
      polviews == 4 ~ "neutral",
      polviews >= 5 ~ "conservative"
    )),
    polviews_new = relevel(polviews_new, ref = "neutral")
  ) |> 
  select(numtot, sex_new, sexornt_new, polviews_new, age, degree, numwomen, nummen) |> 
  drop_na() |> 
  filter(numwomen <= 989 & nummen <= 989) # Drop NA

# Model
numtot_mod1 <- glm(formula = numtot ~ sex_new + sexornt_new, data = df, family = "poisson")
numtot_mod2 <- glm(formula = numtot ~ sex_new + sexornt_new + age, data = df, family = "poisson")
numtot_mod3 <- glm(formula = numtot ~ sex_new + sexornt_new + age + degree, data = df, family = "poisson")
numtot_mod4 <- glm(formula = numtot ~ sex_new : sexornt_new + age + degree, data = df, family = "poisson")
msummary(list(numtot_mod1, numtot_mod2, numtot_mod3, numtot_mod4), output = "gt") |> 
  opt_table_font(font = "Optima")
```

The fourth model with interaction term of sex and sex orientation has the lowest AIC, but the third model without interaction term has the lowest BIC.

Interpretation of $\beta$ before the sex = female : sexornt = homo in model 4: Compared to the reference group (male homosexual), female homosexual people have $e^{-0.797}=45.1$ % fewer sexual partners.

## 11.2.3 `sibs`

```{r}
df <- d |> 
  mutate(
    f_hs = ifelse(padeg == 0, F, T),
    m_hs = ifelse(madeg == 0, F, T),
    female = ifelse(sex == 2, T, F),
    race_new = as.factor(case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Other"
    ))
  ) |>
  select(sibs, f_hs, m_hs, female, race_new) |> 
  drop_na()
sibs_mod1 <- glm(formula = sibs ~ f_hs + m_hs, data = df, family = "poisson")
sibs_mod2 <- glm(formula = sibs ~ f_hs + m_hs + female, data = df, family = "poisson")
sibs_mod3 <- glm(formula = sibs ~ f_hs + m_hs + female + race_new, data = df, family = "poisson")
msummary(list(sibs_mod1, sibs_mod2, sibs_mod3), output = "gt") |> 
  opt_table_font(font = "Optima")
```

Model 3 has the lowest AIC and BIC.

The interpretation of $\beta$ before race =white: Compared to the reference group of black people, white people have $e^{-0.437}=64.6$ % fewer siblings.
